---
title: "HW3"
author: "Sarah Munro"
date: "10/13/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(p8105.datasets)
```

#Problem 1

_Load the dataset instacart_
```{r}
library(p8105.datasets)
data("instacart")
```

```{r}
view(instacart)
```

_Describe the dataset instacart_

The dataset `instacart` has `r nrow(instacart)` observations and `r ncol(instacart)` variables. The dataset contains information on products ordered from a grocery store with key variables including `order_id` `aisle` `product_name` and `department`. The variables provide a very detailed breakdown of the orders describing the order in which products were added as well as the hour of the day the order was placed. 

_Count the number of aisles and the aisles with the most items orderd_
```{r}
count(instacart, aisle, name = "n_orders") %>%
arrange(desc(n_orders)) 
##There are 134 asiles in total. The aisles for fresh vegetables and fresh fruits received the most orders, with 150,609 items and 150,473 items ordered respectively.  
```

_Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered_
```{r}
instacart %>%
  count(aisle, name = "n_orders") %>%
  filter (n_orders > 10000) %>%
  ggplot(aes(x=aisle, y=n_orders)) + 
  geom_point(color = "magenta") + 
  theme(
    axis.text.x = element_text(angle=90, hjust=1)
  ) +
  labs(title = "Quantity of Items Ordered by Aisle", x = "Aisle Name", y = "Number of Items Ordered") 
 
```

_Create a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”_ 
```{r}
instacart %>%
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>%
  group_by(aisle, product_name) %>%
  summarize(order_frequency = n()) %>%
  top_n(3) %>%
  arrange(desc(order_frequency)) %>%
  knitr::kable ()
``` 

_Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week_
```{r}
instacart %>%
  filter( product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>%
  group_by(product_name, order_dow) %>%
  summarize(avg_hour = mean(order_hour_of_day)) %>%
  select(product_name, order_dow, avg_hour) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = avg_hour) %>%
knitr::kable ()
```

#Problem 2

_Clean the dataset BRFSS_
```{r}
data("brfss_smart2010")
brfss = brfss_smart2010 %>%
  janitor::clean_names() %>% 
  filter(topic == "Overall Health", response == "Excellent" | response == "Very Good" | response == "Good" | response == "Fair" | response == "Poor") %>%
mutate(response = factor(response, levels =  c("Poor","Fair","Good","Very Good","Excellent")))
  brfss
```

_In 2002, which states were observed at 7 or more locations? What about in 2010?_
```{r}
brfss %>%
  filter(year == 2002 ) %>%
  group_by(locationabbr) %>%
  summarize(n_locations = n_distinct(geo_location)) %>%
  select(locationabbr, n_locations) %>%
  arrange(desc(n_locations))
## There are 6 states (PA, MA, NJ, CT, FL, NC) that were observed at 7 or more locations in 2002
```
```{r}
brfss %>%
  filter(year == 2010 ) %>%
  group_by(locationabbr) %>%
  summarize(n_locations = n_distinct(geo_location)) %>%
  select(locationabbr, n_locations) %>%
  arrange(desc(n_locations))
## In 2010 there were 14 states observed at 7 or more locations
```

_Make a new dataset and a spaghetti plot showing the average data values over time for each state_
```{r}
excellent_health = brfss %>%
  filter(response == "Excellent") %>%
  group_by(year, locationabbr) %>%
  summarize(avg = mean(data_value)) %>%
  select (year, locationabbr, avg) %>%
ggplot(aes(x=year, y=avg)) +geom_line(aes(color=locationabbr)) + labs(title = "Average State Data Values Across Time", y = "Average Data Values")
excellent_health
```

_Make a two-panel plot showing distribution of data_value for responses among locations in NY State_
```{r}
brfss %>%
  filter (locationabbr == "NY", year == 2006 | year == 2010) %>%
  ggplot(aes(x=locationdesc, y=data_value)) + geom_point(aes(color=response)) + facet_grid(. ~year) + theme(
    axis.text.x = element_text(angle=90, hjust=1)) +
  labs(title = "Distribution of Data Value Among NY")
```  

#Problem 3

_Load and tidy the new dataset_
```{r}
acceleration = read_csv("./Data/accel_data.csv") %>%
  janitor::clean_names() %>%
  drop_na()%>%
  mutate(time_of_week = if_else(day == "Saturday" | day == "Sunday", "weekend", "week_day")) %>%
select(week, day_id, day, time_of_week, everything())
acceleration
```
The dataset `acceleration` has `r nrow(acceleration)` observations and `r ncol(acceleration)` variables. Key variables include `week` `day_id` and `time_of_week`. The activity variables correspond to the average activity for each minute of the day.

_Aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals_
```{r}
acceleration %>%
    pivot_longer(activity_1:activity_1440,
    names_to = "minute",
    values_to = "activity"
    ) %>%
    group_by(week, day_id, day, time_of_week) %>%
    summarize(total_activity = sum(activity)) %>%
knitr::kable()
```
_There appears to be no noticeable trends, he is not consistently more or less active on weekdays or weekends. It seems that he took his wrist band off on the 24th and 31st day of the month, which are both saturdays._ 

_Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week_
```{r}
acceleration %>%
  pivot_longer(activity_1:activity_1440,
               names_to = "minute",
               values_to = "activity") %>%
separate(minute, into = c("activity_name", "minute")) %>%
mutate(minute = as.numeric(minute)) %>%
  ggplot(aes(x=minute, y=activity)) + geom_point(aes(color=day)) + labs(title="Activity Breakdown by Day", x = "Hour of the Day") +scale_x_continuous(
    breaks = c(1, 60, 120, 180, 240, 300, 360, 420, 480, 540, 600, 660, 720, 780, 840, 900, 960, 1020, 1080, 1140, 1200, 1260, 1320, 1380, 1440),
    labels = c("12 AM", "1 AM", "2 AM", "3 AM", "4 AM", "5 AM", "6 AM", "7 AM", "8 AM", "9 AM","10 AM", "11 AM", "12 PM", "1 PM", "2 PM", "3 PM", "4 PM", "5 PM", "6 PM", "7 PM", "8 PM", "9 PM", "10 PM", "11 PM", "12 AM")) + theme(axis.text.x = element_text(angle=50, hjust=1)) + 
  viridis::scale_color_viridis(
    name = "day", 
    discrete = TRUE
  )
```  
_There are not many discernable patterns from the plot. One can see that he is relatively unactive, with most measurements falling below 2500. Perhaps he works a desk job where he doesn't get a lot of movement. His activity peaks on Fridays between 8-10 pm, and on wednesdays around 8 pm. Sunday seems to have the highest midday peaks. Activity is very low for all days between 10 pm and 5 am, which is to be expected while he is sleeping._
